---
output: github_document
bibliography: "./inst/REFERENCES.bib"
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures"
)
# Please put your title here to include it in the file below.
Title <- "Title of your paper goes here"
```

# elco

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
<!-- badges: end -->

elco provides functions to handle and (rudimentary) analyse element content data and isotope signatur data. Features are:

1. **Import of XRF and IRMS data** as they can be exported from X-ray fluorescence spectrometers (Rigaku ZSX Primus II) and EA-IRMS devices (HEKATech's Eurovector
EA3000 coupled with Nu Instruments Nu Horizon).

2. Graphical display for easy **data checking**.

3. **Correction of IRMS-measured element content data and isotope signatures** in cases of a mismatch in the signal areas between samples and standards.

4. **Unit conversion** of element content data (e.g. g/g to mol).

5. Computation of the **nominal oxidation state of carbon (NOSC), oxidative ratio (OR), and degree of unsaturation (DU)** [@Worrall.2016b].

```{r, out.width="50%", fig.align='center', echo=FALSE}
knitr::include_graphics("man/figures/elco-features.svg")
```

### How to install

elco can be installed from GitHub:

```{r, eval = FALSE}
remotes::install_github("henningte/elco")
```

### How to use

Here is a short overview on elco general-purpose capabilities (handling element content ratios). Examples of how to import and correct data are not described here, but in the vignettes.

Required packages
```{r, mesage = FALSE, warning=FALSE}
library(elco) # load elco
library(dplyr)
library(quantities)
library(magrittr)
```


First, we have a short look at the sample data and its structure:
```{r}
d <- elco::chno
d
```

Contents for one element are stored in a numeric vector. Measurement units and errors are tracked using the [quantities](https://github.com/r-quantities/quantities) package. elco adds to this an identifier for the element:

```{r}
# show exemplary structure for the C content
d$C
```

New elco objects can be defined with `elco_new_elco` (by providing a quantities object and an element symbol). For instance, this specifies two new C content values:

```{r}
x <- 
  quantities::set_quantities(c(0.3, 0.43), unit = "g/g", errors = c(0.2, 0.12)) %>%
  elco_new_elco(el_symbol = "C")
x
```

elco supports unit conversion (using the functions of the quantities package and molar masses from [PeriodicTable](https://github.com/cran/PeriodicTable)):
```{r}
# g/g to mol
x %>% 
  elco_elco_convert(to = "mol", 
                    sample_mass = quantities::set_quantities(1, unit = "g", errors = 0))

# g/g to mg
x %>% 
  elco_elco_convert(to = "mg", 
                    sample_mass = quantities::set_quantities(1, unit = "g", errors = 0))
```

... even with complete data.frames!
```{r}
# g/g to mol
d %>% elco_elco_convert_df(to = "mol", sample_mass = d$sample_mass)
```

elco helps computing element ratios:

```{r}
d %>%
  elco_elco_convert_df(to = "mol", sample_mass = d$sample_mass) %>%
  dplyr::mutate(cn_molar = C/N,
                ch_molar = C/H,
                co_molar = C/O)
```

Finally, elco provides functions to compute the nominal oxidation state of carbon (NOSC), oxidative ratio (OR), and degree of unsaturation (DU) [@Worrall.2016b]

```{r}
d <- 
  d %>%
  elco::elco_elco_convert_df(to = "mol", sample_mass = d$sample_mass) %>%
  dplyr::mutate(
    nosc = elco_nosc(C, H, N, O),
    or = elco_or(C, H, N, O),
    du = elco_du(C, H, N)
  )
```

Further information can be found in the vignettes:

+ [Reformatting and checking raw data (csv files) from the X-ray fluorescence device (ZSX Primus II, Rigaku)](vignettes/v001-xrf.Rmd)

+ [Importing and Manipulating IRMS data](vignettes/v002-irms-functions.Rmd)

+ [Computation with element contents - unit conversion, element ratios, nominal oxidation state of carbon, oxidative ratio, and degree of unsaturation](vignettes/v004-element-ratios.Rmd)

### How to cite

Please cite this compendium as:

> Teickner, H. and Knorr, K.-H., (`r format(Sys.Date(), "%Y")`). _elco: Handling data on chemical element contents and isotope signatures._. Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://github.com/henningte/elco>

### Acknowledgments

We thank @Gaka.2019 for the provision of test data.

### Licenses

**Text and figures :**  [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)

**Code :** See the [DESCRIPTION](DESCRIPTION) file

**Data :** [CC-BY-SA-4.0](https://creativecommons.org/licenses/by-sa/4.0/deed.en). All external data (files "xrf1.csv", "xrf2.csv", "irms1.csv", and "irms2.csv") for elco were provided by @Gaka.2019. `elco::irms_standards` was collected from different data sources [@Coplen.2006; @IVAAnalysentechnikGmbHuCoKGt.2016] and own measurements (see the documentation for details). `elco::chno` was created by the package's authors.

### Contributions

We welcome contributions from everyone. Before you get started, please see our [contributor guidelines](CONTRIBUTING.md). Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.

### Sources

All data for this package are from @Gaka.2019.

